---
title: "Dormancy_Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Defining Data Sets

Chloe's Dormancy Data
```{r}
dormant <- read.csv("Dormant_Data.csv")
```

Mutated Genes from https://www.cbioportal.org/study/summary?id=msk_met_2021
```{r}
mutated <- read.csv("Mutated_Genes.csv")
```

Changing the column name for gene to "Gene" in Chloe's dormancy data and innerjoining the two sets by the Gene column
```{r}
colnames(dormant)[4] <- "Gene"
data <- dplyr::inner_join(dormant, mutated, by = "Gene")

```

### Quick numbers regarding genes the two sets share

Number of genes in Chloe's data
```{r}
length(dormant$Gene)
```

Number of genes in cbioportal data
```{r}
length(mutated$Gene)
```

Number of Genes shared between the two

```{r}
length(data$Gene)
```

data will now have genes from the innerjoined set that have data from at least 4 studies in Chloe's data
```{r}
data <- data[data$cell_line_counts >= 4, ]
```

## Upregulated Genes
Upregulated genes will be defined as genes with at least 4 cell line counts and 1 or less downregulated count.
```{r}
upreg <- data[data$downreg_counts <= 1, ]
```

The amount of upregulated genes fitting our description above
```{r}
length(upreg[, 'Gene'])
```

The upregulated genes
```{r}
upreg$Gene
```

Frequencies of upregulated genes from cbioportal data
```{r}
rbind(upreg$Gene, upreg$Freq)
```
CDKN1A and SMAD3 have the highest frequencies in the upregulated genes. I will examine those two a little closer.

### CDKN1A

```{r CDKN1A}
CDKN1A <- upreg[upreg$Gene == "CDKN1A", ]
```

Mean of CDKN1A cell line counts
```{r}
mean(as.numeric(CDKN1A[which(!is.na(CDKN1A[5:18])) + 4]))
```

Median of CDKN1A cell line counts
```{r}
median(as.numeric(CDKN1A[which(!is.na(CDKN1A[5:18])) + 4]))

```
### SMAD3

```{r SMAD3}
SMAD3 <- upreg[upreg$Gene == "SMAD3", ]
```

Mean of SMAD3 cell line counts
```{r}
mean(as.numeric(SMAD3[which(!is.na(SMAD3[5:18])) + 4]))
```
Median of SMAD3 cell line counts
```{r}
median(as.numeric(SMAD3[which(!is.na(SMAD3[5:18])) + 4]))
```


## Downregulated Genes
Downregulated genes will be defined as genes with at least 4 cell line counts and 1 or less upregulated count.
```{r}
downreg <- data[data$upreg_counts <= 1, ]
```

Amount of downregulated genes fitting our description above
```{r}
length(downreg[, 'Gene'])
```

The downregulated genes
```{r}
downreg$Gene
```


Frequencies of downregulated genes

```{r}

rbind(downreg$Gene, downreg$Freq)

```

BRCA2, RB1, and PRBM1 have the highest frequency of mutation in down regulated genes. I will examine those three a little closer

### BRCA2
```{r BRCA2}
BRCA2 <- downreg[downreg$Gene == "BRCA2", ]
```

Mean of BRCA2 cell line counts
```{r}
mean(as.numeric(BRCA2[which(!is.na(BRCA2[5:18])) + 4]))
```

Median of BRCA2 cell line counts
```{r}
median(as.numeric(BRCA2[which(!is.na(BRCA2[5:18])) + 4]))
```

### RB1

```{r RB1}
RB1 <- downreg[downreg$Gene == "RB1", ]
```

Mean of RB1 cell line counts
```{r}
mean(as.numeric(RB1[which(!is.na(RB1[5:18])) + 4]))
```

Median of RB1 cell line counts
```{r}
median(as.numeric(RB1[which(!is.na(RB1[5:18])) + 4]))
```

### PBRM1

```{r PBRM1}
PBRM1 <- downreg[downreg$Gene == "PBRM1", ]
```

Mean of PBRM1 cell line counts
```{r}
mean(as.numeric(PBRM1[which(!is.na(PBRM1[5:18])) + 4]))
```

Median of PBRM1 cell line counts
```{r}
median(as.numeric(PBRM1[which(!is.na(PBRM1[5:18])) + 4]))
```


## Extremely up or down regulated
Now we will look into the genes that are either very upregulated or very donwregulated. We will define these genes by having at least 6 cell line counts with either all upregulated or all downregulated. 
```{r}
temp_data <-dplyr::inner_join(dormant, mutated, by = "Gene")

extr_data <- temp_data[temp_data$cell_line_counts >= 6, ]
extr_up <- extr_data[extr_data$downreg_counts == 0, ]
extr_down <- extr_data[extr_data$upreg_counts == 0, ]

```

### Extreme Up
Frequencies of the very upregulated genes
```{r Extreme Up}

rbind(extr_up$Gene, extr_up$Freq)

```

### Extreme Down
Frequencies of the very downregulated genes
```{r Extreme Down}

rbind(extr_down$Gene, extr_down$Freq)

```

## Genes of Most Interest

### Up: CDKN1A

### Down: BRCA2

## Primary with metastasis vs without 

Defining data sets from https://www.cbioportal.org/study/summary?id=msk_met_2021. prim_met has all of the mutated genes data from samples prom primary tumors with metastasis, while prim_no_met has the mutated genes data from primary tumors with no metastasis
```{r}

prim_met <- read.csv("prim_met.csv")
prim_no_met <- read.csv("prim_no_met.csv")

```

### CDKN1A

Frequency of CDKN1A in primary with metastasis
```{r}
prim_met[prim_met$Gene == "CDKN1A", "Freq"]
```

Frequency of CDKN1A in primary without metastasis
```{r}
prim_no_met[prim_no_met$Gene == "CDKN1A", "Freq"]

```

### BRCA2

Frequency of BRCA2 in primary with metastasis
```{r}
prim_met[prim_met$Gene == "BRCA2", "Freq"]
```

Frequency of BRCA2 in primary without metastasis
```{r}
prim_no_met[prim_no_met$Gene == "BRCA2", "Freq"]
```

Interesting that BRCA2 is mutated more in samples without metastasis